The effects of context inference on motor learning: savings, de-adaptation and spontaneous recovery

### 2021-08-31 adaptation/article/article.tex
Finished intro.

\section{Introduction}
It has been long known that humans are capable of learning multiple novel motor tasks at the same time, even when these tasks oppose each other \cite[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Functional_1997,}, although interference can occur between tasks \cite[e.g.][]{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}.

In order to study motor adaptation, motor error is defined as the deviation between the observed movements of a participant and a baseline movement. For example, in experiments with a mechanical arm \cite[e.g.][]{Gandolfo_Motor_1996}, a correlation coefficient is defined between the movements when the mechanical arm exerts a force and the movements when it does not; the higher the correlation coefficient, the lower the error. Other examples include measuring the force profile in error-clamp trials \cite{Smith_Interacting_2006}, endpoint errors in visuomotor rotations \cite{Kim_Neural_2015} and saccade adaptation \cite{Catz_Cerebellardependent_2008}.

@@

In order to explain these phenomena, a number of computational models have been introduced. The least complex of these models are linear learners [CITATION NEEDED], in which a single one-dimensional variable is used to describe the adaptation of motor commands given the observed motor errors. By having multiple time scales, each with a learning rate and a forgetting rate, these models have been fitted to experimental data, explaining these distinct experimental phenomena.

Bayesian accounts of motor adaptation have also been previously presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between internal models of adaptation via prediction error generated by these internal models. These models replace the idea of explicit multiple time scales with an estimation of their own uncertainty: as the model practices the same task, the uncertainty on its estimate of the underlying modification is reduced, which lowers variability in responses, as well as increases the stability of the estimate.

While these general models of adaptation explain the most common phenomena observed in experiments, a number of known phenomena remain outside of the scope of these models. (1)

[unexplained phenomena. The truth is out there. I want to believe]

@@

In what follows, we introduce the different parts of the model and explain how they work together. In addition, we will highlight experimental findings in motor adaptation from previous works for which our model provides explanations that are otherwise unavailable.

@@

We now briefly describe the three components of the model separately. For clarity, we start with the second component (i.e. motor adaptation), in order to introduce the terminology that we use throughout this work. For a complete mathematical description of the model, see the Methods section.

@@

It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \cite{Wolpert_Multiple_1998a} \todo{Is this the original one? Include more}. While a number of models for motor adaptation exist [EGs], we chose to make use of an exact Bayesian learner. As we discuss below, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminates the need for explicit multiple time scales of learning. We show that this adaptation mechanism can deal with adaptation after any number of trials, ranging from a handful of trials in an experiment to the level of expertise a professional dancer has accrued through years of practicing the same movements.

To describe this component of the model, let us return to reaching movements. Throughout our lives, we have learned the equivalence between a motor command and its outcomes in the form of a forward model. We can write this forward model as follows:
\begin{equation}
p(\vartheta | s, c) = f(s, c; \beta)
\end{equation}
where $\vartheta$ are the outcomes of an action (in the example, direction and velocity of the reach movement), $s$ is the current state (e.g. the current position of the hand), $c$ is a motor command and $\beta$ are the parameters of the forward model $f(\cdot)$. Throughout our lives, the parameters $\beta$ have been fine tuned to produce an accurate forward model $f$ to guide our movements and it is these parameters which can be updated during motor adaptation.

When an observed outcome does not match the predictions produced by $f$, the brain uses the prediction error to update the parameters $\beta$ of the forward model. To describe this adaptation process, our model makes uses of Bayesian inference, such that:
\begin{equation}
q(\beta | \vartheta, s, c) \propto p(\vartheta | s, c, \beta)p(\beta)
\end{equation}
where $q(\beta | ...)$ represents the new estimate of $\beta$ after having observed the outcome $\vartheta$ of the previous motor command $c$. $p(\vartheta | s, c, \beta)$ is the \textit{a priori} probability of observing the outcome $\vartheta$ as predicted by the forward model $f$, and $p(\beta)$ is the previous estimate of $\beta$.

\subsection{Context inference}
In our model, the idea of context is quite general. The context comprises the relevant elements of the environment (e.g. the strength and direction of the wind when walking) and of the task at hand (immediate and long-term goals, reward structure, etc). Previous models \cite[e.g.][]{Wolpert_Multiple_1998a,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g. $f_B$ and $f_H$ from the previous section) make different predictions for the outcome of that action; the model that best predicts that action is deemed to be the ``correct'' model, i.e. the model that best represents the context.

In this work, we expand on this idea of switching internal models by creating context inference that integrates several sources of information, including sensory cues, outcome prediction errors and reward prediction errors.  We show how the incorporation of all available contextual information explains many previously-unexplained phenomena from experiments with human subjects, without invoking \textit{ad hoc} model components.

@@

In the previous section, we used the variable $\vartheta$ to refer to the outcomes of actions. In this section, we expand the definition of an observation $\vartheta$ to include any information given to the decision-making agent by the environment. This includes, as before, the outcomes of actions (e.g. the new position of the hand), but also any other sensory information that might be indirectly related to actions, or even completely intependent from them. For this section, the most important component of $\vartheta$ is sensory contextual information, i.e. information that might directly help infer the identity of the current context.

In experimental settings, the sensory contextual information can take the form of a visual cue \cite[e.g.][]{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \cite[e.g.][]{Forano_Timescales_2020} or even the start of a new block of trials \cite{Ethier_Spontaneous_2008}. All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta$ form part of $p(\vartheta_t | \vartheta_{t-1}, c_{t-1})$ and are integrated together to identify a context.

@@

\subsection{Action selection}
Action selection is affected by the current context via the context-dependent forward model that is used. In our model, action selection is made by building a distribution over available actions which is a weighted sum of the distributions given by the existing forward models, where the weights are obtained by the context inference component of the model. In other words, a distribution is created as follows:
\begin{equation}
p(c_t) \propto \displaystyle\sum_{\zeta \in \Phi}q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) p(c_t | \zeta)
\end{equation}
From this distribution, the current action (motor command $c_t$) can be sampled and carried out.

@@

\subsection{Savings and quick de-adaptation}
Savings refers to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it. Savings is almost universally observed in human participants [A THOUSAND CITATIONS, c.f. papers with volatility or ramping up/down], but animal experiments have shown mixed results \citep[e.g.][]{Kojima_Memory_2004}. In an O-A-O-A experiment, for example, savings would be seen in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block.

On the other hand, quick de-adaptation refers to the observation that in an O-A-O paradigm, the adaptation from O to A is much slower than that from A back to O.

In this section, we will show that both phenomena are the two sides of the same coin, and the speed at which this quick adaptation/de-adaptation happens, which is known not to be immediate [Citations] is related to the availability and reliability of contextual information, as well as to the quality of the prediction errors.

To show this, we categorized experiments performed with human participants by looking at the amount of contextual information that made available to participants. In some experiments [EGssss], the context is clearly revealed to the participant. We call these cued-context experiments. In other experiments, partial information is available to participants [EGsss] in the form of prediction errors, partial contextual information or reward prediction errors; we refer to these as partially-cued experiments. While there are no experiments in which no contextual information is available, there are some experiments which contain blocks of trials in which no contextual information is available and participants have no way of inferring the current context. Abusing our notation, we call these experiments (or rather, these sections of the experiments) uncued experiments. In this section, we focus on the two former categories, while the latter is left for the next sections.

Note that, because context inference integrates information from many sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use. For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone. In fact, the mere act of holding the handle informs participants that something is different from baseline. Proprioceptive signals can also provide contextual information [EGs of experiments with hidden hands]. The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than normal trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite[e.g.][]{Ethier_Spontaneous_2008}.

@@

Note that the data from \cite{Oh_Minimizing_2019} and \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism in our model. \todo{Need to be careful here, as de-adaptation is sometimes discussed only in terms of experiments with no contextual cues or where the meaning of the cues is inverted}.

@@

\subsection{Action selection in error-clamp blocks}
During error-clamp blocks at the end of block sequences, participants' behavior can be described as the display of spontaneous recovery (i.e. displaying adaptation consistent with a previously-encountered context) during the early trials of the E block, followed by a slow return to baseline across as many as hundreds of trials. However, the direction of the spontaneous recovery, its length, the delay before it is observed, the speed of the return to baseline and the final asymptote of the adaptation seem to vary greatly depending on the experiment [EGs].

@@

Similar models have been presented before \cite[e.g.][]{Wolpert_Multiple_1998a,Baddeley_System_2003}, which infer the internal models (forward and backward) that best explain the observations and use them to issue motor command. We built on this studies in two important ways: (1) we generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task; (2) we showed how this seemingly small generalization provides a parsimonious explanation for many experimental phenomena across many experiments in motor adaptation that, up to now, had required multiple \textit{ad hoc} mechanisms.

@@

It is important to note that the most essential part of the model we presented is the concept of the context and the inclusion of a Bayesian inference scheme over the context, which integrates all the information that is available to the participant during the experiment. The mathematical details of the model are not strictly necessary to reproduce the results. Most notably, we chose to make use of exact Bayesian inference for motor adaptation, with conjugate priors to derive exact update equations. This choice greatly simplifies our simulations and provides very intuitive intepretations for the results. However, we expect the results we presented here to hold using alternative models for motor adaptation, such as Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998a,Kording_Bayesian_2004}. A well established family of models for motor adaptation is linear learners \cite[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Lee_Dual_2009}, however, these models do not incorporate uncertainty in the estimate of the parameters of the internal models, which would limit their implementation in our hierarchical model.

@@

The most direct secondary contextual information comes in the form of reward and end-point feedback. Participants are told whether they obtained the desired reward at the end of a trial and are shown the end point of their movement. When a participant observes an error that is too large to explain with the expected motor variability, this information can be used to infer that the inferred context might be wrong. This is the case of the experiments by \cite{Oh_Minimizing_2019} shown in \fref{fig:oh-2019}B-C: if the adaptation is high, changes in context produce errors much larger than those of motor variability, and a context switch is easily and immediately identified.; if adaptation is low, the errors produced by context switching are closer in magnitude to motor variability and evidence accumulation is necessary.

@@

This model is a Bayesian alternative to the N-state learners from
\citep[e.g.][]{Lee_Dual_2009}. In this model, there are no explicit different
timescales and multiple states; instead, the model relies on Bayesian learning
for motor adaptation, combined with inference over the current context.

@@

Context inference can work through context-specific (possibly unreliable) cues,
or through the dynamics of the system (e.g. expected outputs of motor
commands), using prediction error as the force behind inference. In this case,
the generative model can is written as follows:
\[
p(z_t, z_{t-1}, a_{t-1}, \theta_t, \theta_{t-1}) = p(\theta_t | z_t)p(z_t|\theta_{t-1}, a_{t-1})
\]
and inferring the current (hidden) generalized state is done with Bayes' theorem:
\[
  q(z_t | \theta_t, \theta_{t-1}, a_{t-1}) \propto p(\theta_t | z_t)p(z_t|\theta_{t-1}, a_{t-1})
\]

@@

In the following sections, we present a specific mathematical implementation
of the principles outlined above. This is the implementation used throughout the
simulations in the main text.

@@

We do not directly model hand movements or eye position. Instead, we model
systems in the motor error space: the position $s_t = 0$ represents a trial in
which no motor error occurred. We make use of the fact that in most
experiments, motor adaptation needs to happen in one of two directions
(e.g. clockwise vs. counter-clockwise) to keep the error space
two-dimensional. With this, $s_t > 0$ represents motor error in one direction
(e.g. clockwise) and $s_t < 0$, error in the other direction. We made this
choice because, in this work, we would model different experiments with
different state spaces; but all these experiments have the commonality of
two-dimensional error spaces. In order to model a specific experiment in its
own physical space (e.g. hand positions in a reaching task), only the dynamical
system would need to be adapted.

We assume that the decision-making agent has an internal forward model that
maps motor commands to outcomes:
\begin{equation}
p(s_{t+1} | s_t, a_t, m) = f_m(s_t, a_t)
\end{equation}
where $a_t$ are motor commands (actions). We further assume that participants
have learned $f_m$ throughout their lives and use it to decide which motor
command to issue. When a perturbation is introduced, to which the participant
must adapt, we assume they learn a new forward model $f_m^{\phi}$, where $\phi$
represents the parameters of the function that participants must learn. For
example, in a saccade task, $\phi$ represents the new gain. Note that $\phi$
depends on the context and participants need to use and update multiple
forward adapted models $f_m^{\phi}$ throughout one experiment.

Given a forward model $f_m$ (or $f_m^\phi$), action selection is done by
building a probability distribution over possible actions $a_t$ based on how
likely they are to produce zero motor error. For example, in saccade
experiments, motor commands are issued based on how likely they are to move the
fovea to the target. These motor commands are issued to counteract intrinsic
sources of error, such as muscle variability \todo{expand this?}, and extrinsic
sources of error, such as the experimental manipulations on saccade gain. Thus,
the expected observed error can be written as:
\begin{equation}
p(\epsilon | a_t) \propto p(\epsilon_{\text{intrinsic}} | a_t) + p(\epsilon_{\text{extrinsic}} | a_t)
\end{equation}
and a motor command can be chosen to maximize the chance of observing zero error:
\begin{equation}
p(a_t) = argmin_{a_t} p(\epsilon | a_t)
\end{equation}

For simplicity, we assume that motor errors are Gaussian, such that:
\begin{equation}
p(\epsilon | a_t) = N(\mu_{\text{extrinsic}}, \sigma_\epsilon)
\end{equation}
where $\sigma_\epsilon$ incorporates both intrinsic and extrinsic sources of
error. The mean of the error is assumed to have no component, as motor
commands, in the absence of external errors, should produce an error centered
at zero [CITATION MISSING].

@@

To update $\phi$, we assumed that for each context, the agent's belief over
the magnitude of the adaptation is given by a normal distribution:
\begin{equation}
p(\phi) = N(\mu_\phi, \sigma_\phi) \label{eqn:data-dist}
\end{equation}
where $\mu_\phi$ and $\sigma_\phi$ are parameters to be estimated at each
trial. Before the experiment begins, the agent will have a prior distribution
over these parameters. A standard Bayesian approach is to choose NormalGamma
priors:
\begin{equation}
p(\mu_\phi, \sigma_\phi) = NG(\mu_0, \nu_0, \alpha_0, \beta_0)
\end{equation}
where $\mu_0, \nu_0, \alpha_0$ and $\beta_0$ are free hyperparameters of the
model. We assume the following hyperparameter values \todo{Justify}:
\begin{align} \mu_f^{(0)} &= (0, -1, 1) \\
  \nu_f^{(0)} &= 0.1 \\
  \alpha_f^{(0)} &= 1 \\
  \beta_f^{(0)} &= 0.5
\end{align}
Note that $\mu_f^{(0)}$ is different for each context and is thus provided as a
vector $(0, -1, 1)$ representing the baseline and two opposing adaptations
(e.g. clockwise and counter-clockwise visuomotor rotations), respectively. For
all other hyperparameters, all contexts have the same value. These values
ensure that the prior over the hyperparameters $(\mu_\phi, \sigma_\phi)$ has the mode
at $(\mu_\phi^{(0)}, 1)$.

@@

The effect of the hyperparameters of the priors are worth a note or two, as
they are complex.

Naturally, $\mu$ affects the initial estimate of the adaptation, in the same
units as the necessary adaptation. $\nu$ encodes how stable this hyperprior is:
higher values (e.g. 10000) all but guarantee that the hypermean will not move
in the face of evidence; In principle, enough evidence should still move it,
but that won't happen during an experiment. Smaller values (e.g. 1 / force\_sd,
as is the default on the code) make the hypermean (and thus the mean) follow
evidence more freely. Note that as more observations are accumulated, $\nu$
becames bigger and bigger, solidifying the value of the hypermean.

>>>>

\section{Introduction}
It has been long known that humans are capable of learning multiple novel motor tasks at the same time, even when these tasks oppose each other \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference can occur between tasks \citep[e.g.][]{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}.

In order to study motor adaptation, motor error is defined as the deviation between the observed movements of a participant and a baseline movement. For example, in experiments with a mechanical arm \citep[e.g.][]{Gandolfo_Motor_1996}, a correlation coefficient is defined between the movements when the mechanical arm exerts a force and the movements when it does not; the higher the correlation coefficient, the lower the error. Other examples include measuring the force profile in error-clamp trials \citep{Smith_Interacting_2006}, endpoint errors in visuomotor rotations \citep{Kim_Neural_2015} and saccade adaptation \citep{Catz_Cerebellardependent_2008}.

@@

In order to explain these phenomena, a number of computational models have been introduced. The least complex of these models are linear learners \citep[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}, in which a single one-dimensional variable is used to describe the adaptation of motor commands given the observed motor errors. By having multiple time scales, each with a learning rate and a forgetting rate, these models have been fitted to experimental data, explaining these distinct experimental phenomena.

Bayesian accounts of motor adaptation have also been previously presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between internal models of adaptation via predictions generated by these internal models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}. These models replace the idea of explicit multiple time scales with an estimation of their own uncertainty: as the model practices the same task, the uncertainty on its estimate of the underlying modification is reduced, which lowers variability in responses, as well as increases the stability of the estimate.

While these general models of adaptation explain the most common phenomena observed in experiments, a number of known phenomena remain outside of their scope. For example, it is known that adaptation rate is reduced in situations where the environment is unstable and unpredictable \citep{Herzfeld_memory_2014}, or situations in which errors are small \citep{Marko_Sensitivity_2012} and slowly introduced \citep{Huang_Persistence_2009}. Learning new adaptations has been found to depend on the history of adaptations learned \citep{Vaswani_Decay_2013}, as well as the way in which those adaptations were presented [CITATION NEEDED].

In this work, we present a computational model for motor adaptation inspired by models such as MOSAIC \citep{Wolpert_Multiple_1998} and Bayesian switching models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}. We expand on these models by expanding the concept of context and introducing an explicit mechanism for inferring the context based on Bayesian inference. In our model, the context encompasses all the information relevant to making decisions, both environmental and internal to the agent making these decisions. To infer the current context, our model integrates several sources of information, including the predictions made by its internal forward models, an estimate of the reliability of these predictions, direct sensory information (such as visual contextual cues), proprioceptive information, and feedback provided by the environment. Having inferred the context, the model can choose the best-fitting internal model with which to make decisions.

We show that the dynamics of context inference provide a unifying explanation for all the experimental phenomena outlined above, providing a single account for motor adaptation under changing contexts, while relying on known mechanisms for adaptation.

@@

We used our model to explain several experimental phenomena found in previous studies. We simulated representative experiments from these studies to show how our model explains their findings using the dynamics of context inference. In what follows, we will present these simulations alongside the experimental data from the representative studies and discuss how context inference explains the data.

Before presenting our simulations, we will briefly explain the fundamentals of the model we present. We leave the mathematical details of the implementation to the Methods section.

@@

We now briefly describe the three components of the model separately. For clarity, we start with the second component (i.e. motor adaptation), in order to introduce the terminology that we use throughout this work.

@@

It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \citep{Wolpert_Multiple_1998}. We chose to make use of an exact Bayesian learner. As we discuss below, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminates the need for explicit multiple time scales of learning. We show that this adaptation mechanism can deal with adaptation after any number of trials, ranging from a handful of trials in an experiment to the level of expertise a professional dancer has accrued through years of practicing the same movements.

To describe this component of the model, we make use of the example of reaching movements. Throughout our lives, we have learned the equivalence between a motor command and its outcomes in the form of a forward model. We can write this forward model as follows:
\begin{equation}
p(\vartheta | s, c) = f(s, c; \phi) \label{eqn:forward-model}
\end{equation}
where $\vartheta$ are the outcomes of an action (in the example, direction and velocity of the hand during the reaching movement), $s$ is the current state (e.g. the current position of the hand), $c$ is a motor command and $\phi$ are the parameters of the forward model $f(\cdot)$. Throughout our lives, the parameters $\phi$ have been fine tuned to produce an accurate forward model $f$ to guide our movements and it is these parameters which are updated during motor adaptation.

When an observed outcome does not match the predictions produced by $f$, the prediction error is used to update the parameters $\phi$ of the forward model. To describe this adaptation process, our model makes uses of Bayesian inference, such that:
\begin{equation}
q(\phi | \vartheta, s, c) \propto p(\vartheta | s, c, \phi)p(\phi)
\end{equation}
where $q(\phi | ...)$ represents the new estimate of $\phi$ after having observed the outcome $\vartheta$ of the previous motor command $c$. The likelihood $p(\vartheta | s, c, \phi)$ is the \textit{a priori} probability of observing the outcome $\vartheta$ as predicted by the forward model $f$, and $p(\phi)$ is the previous estimate of $\phi$.

\subsection{Context inference}
In our model, the idea of context is quite general. The context comprises the relevant elements of the environment (e.g. the strength and direction of the wind when walking) and of the task at hand (immediate and long-term goals, reward structure, etc). Previous models \cite[e.g.][]{Wolpert_Multiple_1998,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g. two forward models $f_B$ and $f_H$) make different predictions for the outcome of that action; the model that best predicts that action is deemed to be the ``correct'' model, i.e. the model that best represents the context.

In this work, we expand on this idea of switching internal models by creating context inference that integrates several sources of information, including sensory cues, outcome prediction errors and reward prediction errors.

@@

In the previous section, we used the variable $\vartheta$ to refer to the outcomes of actions. In this section, we expand the definition of an observation $\vartheta$ to include any information given to the decision-making agent by the environment. This includes, as before, the outcomes of actions (e.g. the new position of the hand), but also any other sensory information that might be indirectly related to actions, or even completely independent from them. For this section, the most important component of $\vartheta$ is contextual sensory information, i.e. information that might directly help infer the identity of the current context.

In experimental settings, the contextual sensory information can take the form of a visual cue \cite[e.g.][]{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \cite[e.g.][]{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or even the start of a new block of trials \cite{Ethier_Spontaneous_2008}. All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta)$ form part of $p(\vartheta_t | \vartheta_{t-1}, c_{t-1})$ and are integrated together to identify a context.

@@

\subsection{Action selection}
Action selection is affected by the current context via the context-dependent forward model that is used. In our model, action selection is made by building a distribution over available actions which is a weighted sum of the distributions given by the existing forward models, where the weights are obtained by the context inference component of the model. In other words, a distribution is created as follows:
\begin{equation}
p(c_t) \propto \displaystyle\sum_{\zeta \in \Phi}q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) p(c_t | \zeta)
\end{equation}
where $p(c_t | \zeta)$ corresponds to \eref{eqn:forward-model}. From this distribution, the current action (motor command $c_t$) can be sampled and carried out.

@@

\subsection{Savings}
Savings refers to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it. Savings is almost universally observed in human participants \cite{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}. In an O-A-O-A experiment, for example, savings would be seen in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block.

In this section, we discuss savings, and the related phenomenon of deadaptation, in terms of the switching between forward models. This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019}, but we show that expanding switching into context inference explains facets of savings that these previous models do not.

To show this, we categorized experiments by looking at the amount of contextual information made available to participants. In some experiments \citep[e.g.][]{Kim_Neural_2015,Lee_Dual_2009}, the context is clearly revealed to the participant. We call these cued-context experiments. In other experiments, partial information is available to participants \citep[e.g.][]{Davidson_Scaling_2004,Zarahn_Explaining_2008} in the form of large prediction errors, partial contextual information or reward prediction errors; we refer to these as partially-cued experiments.

To our knowledge, there are no experiments in which no contextual information is available. However, many experiments \citep[e.g.][]{Huang_Persistence_2009,Brennan_Decay_2015,Smith_Interacting_2006} contain blocks of error-clamp trials in which participants have no way of identifying the current context. These blocks could be considered uncued, however, as we discuss in the following sections, contextual information is still available to participants.

In this section, we focus on cued and partially-cued experiments and leave error-clamp trials to the following sections.

Note that, because context inference integrates information from many sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use. For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone. In fact, the mere act of holding the handle informs participants that something is different from baseline. Proprioceptive signals can also provide contextual information \cite{Dizio_Motor_1995,Shadmehr_Adaptive_1994}. The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than normal trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite[e.g.][]{Ethier_Spontaneous_2008}.

@@

Note that the data from \cite{Oh_Minimizing_2019} and \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism in our model.

@@

\subsection{Action selection in error-clamp blocks}
During error-clamp blocks at the end of block sequences, participants' behavior can be described as the display of spontaneous recovery (i.e. displaying adaptation consistent with a previously-encountered context) during the early trials of the E block, followed by a slow return to baseline across as many as hundreds of trials \citep{Brennan_Decay_2015}. However, the direction of the spontaneous recovery, its length, the delay before it is observed, the speed of the return to baseline and the final asymptote of the adaptation seem to vary greatly depending on the experiment \citep{Brennan_Decay_2015,Vaswani_Decay_2013,Smith_Interacting_2006,Shmuelof_Overcoming_2012}.

@@

Similar models have been presented before \cite[e.g.][]{Wolpert_Multiple_1998,Baddeley_System_2003}, which infer the internal models (forward and backward) that best explain the observations and use them to issue motor command. We built on this studies in two important ways: (1) we generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task; (2) we showed how this seemingly small generalization provides a parsimonious explanation for many experimental phenomena across many experiments in motor adaptation that, up to now, had required multiple \textit{ad hoc} mechanisms.

@@

It is important to note that the most essential part of the model we presented is the concept of the context and the inclusion of a Bayesian inference scheme over the context, which integrates all the information that is available to the participant during the experiment. The mathematical details of the model are not strictly necessary to reproduce the results. Most notably, we chose to make use of exact Bayesian inference for motor adaptation, with conjugate priors to derive exact update equations. This choice greatly simplifies our simulations and provides very intuitive intepretations for the results. However, we expect the results we presented here to hold using alternative models for motor adaptation, such as Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019,Baddeley_System_2003} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004}. A well established family of models for motor adaptation is linear learners \cite[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Lee_Dual_2009}, however, these models do not incorporate uncertainty in the estimate of the parameters of the internal models, which would limit their implementation in our hierarchical model.

@@

The most direct secondary contextual information comes in the form of reward and endpoint feedback. Participants are told whether they obtained the desired reward at the end of a trial and are shown the end point of their movement. When a participant observes an error that is too large to explain with the expected motor variability, this information can be used to infer that the inferred context might be wrong. This is the case of the experiments by \cite{Oh_Minimizing_2019} shown in \fref{fig:oh-2019}B-C: if the adaptation is high, changes in context produce errors much larger than those of motor variability, and a context switch is easily and immediately identified.; if adaptation is low, the errors produced by context switching are closer in magnitude to motor variability and evidence accumulation is necessary.

@@

@@

Context inference can work through context-specific (possibly unreliable) sensory cues,
or through the dynamics of the system (e.g. expected outputs of motor
commands), using prediction error as the force behind inference. In this case,
the generative model can is written as follows:
\[
p(z_t, z_{t-1}, a_{t-1}, \theta_t, \theta_{t-1}) = p(\theta_t | z_t)p(z_t|\theta_{t-1}, a_{t-1})
\]
and inferring the current (hidden) generalized state is done with Bayes' theorem:
\[
  q(z_t | \theta_t, \theta_{t-1}, a_{t-1}) \propto p(\theta_t | z_t)p(z_t|\theta_{t-1}, a_{t-1})
\]

@@

In the following sections, we present a specific mathematical implementation
of the principles outlined above. This is the implementation used throughout the
simulations in the Results section.

@@

We do not directly model hand movements or eye position. Instead, we model
systems in the motor error space: the position $s_t = 0$ represents a trial in
which no motor error occurred. We make use of the fact that in most
experiments, motor adaptation needs to happen in one of two directions
(e.g. clockwise vs. counter-clockwise) to keep the error space
two-dimensional. With this, $s_t > 0$ represents motor error in one direction
(e.g. clockwise) and $s_t < 0$, error in the other direction. We made this
choice because, in this work modeled different experiments with
different state spaces, but with the commonality of (at most)
two-dimensional error spaces. In order to model a specific experiment in its
own physical space (e.g. hand positions in a reaching task), only the dynamical
system would need to be adapted.

We assume that the decision-making agent ha a baseline internal forward models that
map motor commands to outcomes:
\begin{equation}
p(s_{t+1} | s_t, a_t, m) = f_O(s_t, a_t)
\end{equation}
where $a_t$ are motor commands (actions). We further assume that participants
have learned $f_O$ throughout their lives and use it to decide which motor
command to issue under normal circumstances (i.e. no adaptation). When a
perturbation is introduced, to which the participant must adapt, we assume they
learn a new forward model $f_m^{\phi}$, where $\phi$ represents the parameters
of the function that participants must learn. For example, in a saccade task,
$\phi$ represents the new gain. Note that $\phi$ depends on the context and
participants may need to use and update multiple forward adapted models
$f_m^{\phi}$ throughout one experiment.

Given a forward model $f_m$ (or $f_m^\phi$), action selection is done by
building a probability distribution over possible actions $a_t$ based on how
likely they are to produce zero motor error. For example, in saccade
experiments, motor commands are issued based on how likely they are to move the
fovea to the target. These motor commands are issued to counteract intrinsic
sources of error, such as muscle variability and exhaustion, and extrinsic
sources of error, such as the experimental manipulations on saccade gain. Thus,
the expected observed error can be written as:
\begin{equation}
p(\epsilon | a_t) \propto p(\epsilon_{\text{intrinsic}} | a_t) + p(\epsilon_{\text{extrinsic}} | a_t)
\end{equation}
and a motor command can be chosen to maximize the chance of observing zero error:
\begin{equation}
p(a_t) = argmin_{a_t} p(\epsilon | a_t)
\end{equation}

For simplicity, we assume that motor errors are Gaussian, such that:
\begin{equation}
p(\epsilon | a_t) = N(\mu_{\text{extrinsic}}, \sigma_\epsilon)
\end{equation}
where $\sigma_\epsilon$ incorporates both intrinsic and extrinsic sources of
error and is a free parameter of the model. The mean of the error is assumed to
have no intrinsic component, as motor commands, in the absence of external
errors, should produce an error centered at zero [CITATION MISSING].

@@

To update $\phi$, we assumed that for each context, the agent's belief over
the magnitude of the adaptation is given by a normal distribution:
\begin{equation}
p(\phi) = N(\mu_\phi, \sigma_\phi) \label{eqn:data-dist}
\end{equation}
where $\mu_\phi$ and $\sigma_\phi$ are parameters to be estimated at each
trial. Before the experiment begins, the agent will have a prior distribution
over these parameters. A standard Bayesian approach is to choose NormalGamma
priors:
\begin{equation}
p(\mu_\phi, \sigma_\phi) = NG(\mu_0, \nu_0, \alpha_0, \beta_0)
\end{equation}
where $\mu_{\phi}^0, \nu_{\phi}^0, \alpha_{\phi}^0$ and $\beta_{\phi}^0$ are free hyperparameters of the
model.

@@

\subsubsection{Interpreting the hyperparameters}
Naturally, $\mu$ affects the initial estimate of the adaptation, in the same
units as the necessary adaptation. $\nu$ encodes how stable this hyperprior is:
higher values (e.g. 10000) all but guarantee that the hypermean will not move
in the face of evidence; In principle, enough evidence should still move it,
but that won't happen during an experiment. Smaller values (e.g. 1 / force\_sd,
as is the default on the code) make the hypermean (and thus the mean) follow
evidence more freely. Note that as more observations are accumulated, $\nu$
becames bigger and bigger, solidifying the value of the hypermean.

The hyperparameters $\alpha$ and $\beta$ have a more complex effect. Note that the
mean of a Gamma distribution is $1 / (\alpha \nu)$; this mean is being used
as the standard deviation of a Gaussian by the rest of the agent (cheating),
which makes it an important measure of uncertainty. While setting the default
hyperparameters, the values used are $\alpha = 0.5 / \sigma_0$ and
$\beta = 0.5$, where $\sigma_0$ is the \textit{a priori} estimate of the
standard deviation of the force exerted by the environment (force\_sds in the
code), which controls the initial learning rate. This makes the initial
standard deviation equal $\sigma_0$, which makes it consistent with the
fixed-force model in its interface. The 0.5 values ensure that uncertainty is
large at the beginning and during the experiment is greatly reduced, but never
to a point where it's ``visually'' too small. Changing this 0.5 would make the
standard deviation change more quickly, making the model more or less precise
in its predictions, independently of the volatility of the mean of the
adaptation (via the hypermean).
